---
title: "MOT vs Spoken BNC results"
output:
  word_document:
    reference_docx: template.docx
  pdf_document: default

---

Date: `r Sys.Date()`

```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE, include=FALSE}
lib <- c("tidyverse", "magrittr", "knitr", 
         "gridExtra", "grid", "mgcv")
lapply(lib, require, character.only = TRUE)
load(file="ws_report.RData")
```

```{r echo = FALSE}
grid_arrange_shared_legend <- function(..., nrow = 1,y_title, x_size, y_size,widths1 = 2, widths2=0.2,hjust1=1,vjust1=0, x_title, ncol = length(list(...)), position = c("bottom", "right")) {

  plots <- list(...)
  position <- match.arg(position)
  g <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
  lheight <- sum(legend$height)
  lwidth <- sum(legend$width)
  gl <- lapply(plots, function(x) x + theme(legend.position = "none"))
  gl <- c(gl, nrow = nrow, ncol = ncol)

  combined <- switch(position,
                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
                                            legend,
                                            ncol = 1,
                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
                                           legend,
                                           ncol = 2,
                                           widths = c(widths1, widths2),
                                           left=textGrob(y_title, gp=gpar(fontsize=y_size), rot = 90),
                                           bottom=textGrob(x_title, gp=gpar(fontsize=x_size), vjust = vjust1, hjust = hjust1)))
  grid.newpage()
  grid.draw(combined)

}

round_df <- function(df, digits =2) {
  nums <- vapply(df, is.numeric, FUN.VALUE = logical(1))

  df[,nums] <- round(df[,nums], digits = digits)

  (df)
}
```

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "images/"
)
```

```{r echo=FALSE }
BNC_raw_diff <- summary_texts %>%
  (function(x) {
    100 - ((x$Tokens[which(x$Transcription == "BNC-all, Orthographic")] / 
              x$Tokens[which(x$Transcription == "BNC-all-raw, Orthographic")]) * 100)
  }) %>%
  round(2)
```

<br>

NOTE. All words in the complete Spoken BNC (BNC-all-raw, *N* = `r as.character(summary_texts$Tokens[2])`) were phonetically converted. `r BNC_raw_diff`% of words of the BNC-all-raw weren't available in their phonetic forms, resulting in a smaller subset of tokens (BNC-all, *N* = `r as.character(summary_texts$Tokens[3])`). Only BNC-all was used for the following analyses.

<br>

#### Table 1.

Total number of orthographic and phonetic tokens and types in mothers and BNC samples, together with Mean Length of Utterance (MLU) information (NOTE. In the table, BNC-all-raw refers to the raw Spoken BNC which includes `r BNC_raw_diff`% of words not available in their phonetic form. Only BNC-all and BNC-match were used in the following analyses).

```{r echo=FALSE }
summary_texts %>%
  kable()
```

<br>

### Word categories

For word types (see Figure 1 below), plural common nouns and verbs were considered in their root forms ("singular common noun" and "to-infinitive" forms respectively).

<br>

```{r gram_types, echo=FALSE, warning=FALSE}
plot_gram_types <- function(df, title) {
  ggplot(df, aes(x=cat, y=perc))+
  geom_bar(position="dodge",stat="identity")+
  labs(x = "Grammatical categories", y = "Word types (%)") +
  ggtitle(title)+
  theme(axis.text.x = element_text(size=5,angle = 40, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=8))+
  scale_y_continuous(limits = c(0,62.5)) +
  scale_x_discrete(limits = mot_uni_mor_raw$cat)
}

plot1 <- plot_gram_types(mot_uni_mor_raw, "Mothers")
plot2 <- plot_gram_types(bnc_match_ort_types, "BNC-match")
plot3 <- plot_gram_types(bnc_all_ort_types, "BNC-all")

grid.arrange(arrangeGrob(plot1,plot2,plot3,
                         left = textGrob("Word types (%)", rot = 90, vjust = 1),
                         bottom = textGrob("Grammatical categories", rot = 0, vjust = 0.5),
                         nrow = 1))

```

**Figure 1.** Relative word type frequencies of nouns, verbs, adjectives (Adj), adverbs (Adv) and pronouns (Pron) for Mothers, BNC-match and BNC-all. 

<br>

#### Table 2.

Chi-squared analyses comparing the mothers and BNC-match for number of nouns, verbs, adjectives, adverbs and pronouns, for word types (left) and word tokens (right).

```{r echo=FALSE, warning=FALSE}
prop_test <- function(df_mot, df_bnc_match, cat_name, word_type) {
  n_prop_test <- function(df) {
    ( sum(df$freq) / (df$perc %>% sum()) ) * 100
  }
  
  n_mot <- n_prop_test(df_mot)
  n_bnc_match <- n_prop_test(df_bnc_match)
  
  prop_test <- prop.test(x = c(df_mot$freq[which(df_mot$cat == cat_name)], 
                df_bnc_match$freq[which(df_bnc_match$cat == cat_name)]),
                n = c(n_mot, n_bnc_match),
                correct = FALSE)
  
  tibble(!!word_type := cat_name,
         `λ^2` = prop_test$statistic,
         df = prop_test$parameter,
         p = prop_test$p.value)
}

prop_table <- function(df_mot, df_bnc_match, word_type) {
  prop_test(df_mot, df_bnc_match, "NOUN", word_type) %>%
  rbind(prop_test(df_mot, df_bnc_match, "VERB", word_type)) %>%
  rbind(prop_test(df_mot, df_bnc_match, "ADJ", word_type)) %>%
  rbind(prop_test(df_mot, df_bnc_match, "ADV", word_type)) %>%
  rbind(prop_test(df_mot, df_bnc_match, "PRON", word_type)) %>%
  mutate(`λ^2` = round(`λ^2`, 2),
         p = ifelse(p < 0.001, "<0.001", round(p, 2)))
}

prop_table(mot_uni_mor_raw, bnc_match_ort_types, "Word types") %>%
  cbind(prop_table(mot_mor_raw, bnc_match_gram_tokens, "Word tokens")) %>%
  kable()
```

<br>

#### Table

95% confidence intervals for grammatical category proportions (**Word types**) through logistic regression.

```{r echo=FALSE, message=FALSE}
CI_table(mot_uni_mor_raw, bnc_match_ort_types) %>%
  kable()
```

<br>

```{r gram_tokens, echo=FALSE, warning=FALSE}
plot_gram_tokens <- function(df, title) {
  ggplot(df, aes(x=cat, y=perc))+
  geom_bar(position="dodge",stat="identity")+
  labs(x = "Grammatical categories", y = "Word tokens (%)") +
  ggtitle(title)+
  theme(axis.text.x = element_text(size=5,angle = 40, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=8))+
  scale_y_continuous(limits = c(0,32.5)) +
    scale_x_discrete(limits = mot_mor_raw$cat)
}
  
plot1 <- plot_gram_tokens(mot_mor_raw, "Mothers")
plot2 <- plot_gram_tokens(bnc_match_gram_tokens, "BNC-match")
plot3 <- plot_gram_tokens(bnc_all_gram_tokens, "BNC-all")

grid.arrange(arrangeGrob(plot1,plot2,plot3,
                         left = textGrob("Word tokens (%)", rot = 90, vjust = 1),
                         bottom = textGrob("Grammatical categories", rot = 0, vjust = 0.5),
                         nrow = 1))

```

**Figure 2.** Relative word token frequencies of verbs, pronouns (Pron), nouns, adverbs (Adv) and adjectives (Adj) for mothers, BNC-match and BNC-all.

<br>

#### Table

95% confidence intervals for grammatical category proportions (**Word tokens**) through logistic regression.

```{r echo=FALSE, message=FALSE}
CI_table(mot_mor_raw, bnc_match_gram_tokens) %>%
  kable()
```

<br>

### Utterance-final word categories

```{r gram_last_tokens, echo=FALSE, warning=FALSE}
plot_gram_tokens <- function(df, title) {
  ggplot(df, aes(x=cat, y=perc))+
  geom_bar(position="dodge",stat="identity")+
  labs(x = "Grammatical categories", y = "Word tokens (%)") +
  ggtitle(title)+
  theme(axis.text.x = element_text(size=5,angle = 40, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=8))+
  scale_y_continuous(limits = c(0,32.5)) +
    scale_x_discrete(limits = mot_mor_raw_last$cat)
}
  
plot1 <- plot_gram_tokens(mot_mor_raw_last, "Mothers")
plot2 <- plot_gram_tokens(bnc_match_last_gram_tokens, "BNC-match")
plot3 <- plot_gram_tokens(bnc_all_last_gram_tokens, "BNC-all")

grid.arrange(arrangeGrob(plot1,plot2,plot3,
                         left = textGrob("Word tokens (%)", rot = 90, vjust = 1),
                         bottom = textGrob("Grammatical categories", rot = 0, vjust = 0.5),
                         nrow = 1))

```

**Figure 3.** Relative word token frequencies of nouns, pronouns (Pron), verbs, adverbs (Adv) and adjectives (Adj) for mothers, BNC-match and BNC-all.

<br>

#### Table 3. 

Chi-squared analyses comparing the mothers and BNC-match for number of nouns, pronouns, verbs, adverbs and adjectives occurring in utterance-final position. 

```{r echo=FALSE, warning=FALSE}
prop_table(mot_mor_raw_last, bnc_match_last_gram_tokens, "Word tokens") %>%
  slice(c(1,5,2,4,3)) %>%
  kable()
```

<br>

#### Table

95% confidence intervals for grammatical category proportions (**Word tokens**) through logistic regression.

```{r echo=FALSE, message=FALSE}
CI_table(mot_mor_raw_last, bnc_match_last_gram_tokens) %>%
  kable()
```

<br>

### Phonotactic probability and neighbourhood density

For these measures, plural common nouns were considered as singular.

```{r pp_dens, echo=FALSE, warning=FALSE}
plot_pp <- function(df, title) {
  ggplot(df, aes(x=pp))+
  geom_density()+
  scale_y_continuous(limits = c(0,200)) +
  scale_x_continuous(breaks = seq(0,0.03,0.010), limits = c(0,0.03)) +
  labs(x = "Phonotactic probability", y = "Density") +
  ggtitle(title)+
  theme(axis.text.x = element_text(size=6),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=8))
}
  
plot1 <- plot_pp(mot_on_pp_pn, "Mothers")
plot2 <- plot_pp(bnc_match_ph, "BNC-match")
plot3 <- plot_pp(bnc_all_ph, "BNC-all")

grid.arrange(arrangeGrob(plot1,plot2,plot3,
                         left = textGrob("Density", rot = 90, vjust = 1),
                         bottom = textGrob("Phonotactic probability", rot = 0, vjust = 0.5),
                         nrow = 1))
```

**Figure 4.** Density plots indicating the number of word types that occur for each phonotactic probability, for mothers, BNC-match, and BNC-all. 

<br>

```{r pn_dens, echo=FALSE, warning=FALSE}
plot_pn <- function(df, title) {
  ggplot(df, aes(x=pn))+
  geom_bar(aes(y = (..count..)/sum(..count..))) + 
  scale_y_continuous(labels=scales::percent, limits = c(0,0.25)) +
  scale_x_continuous(breaks = seq(0,56,8)) +
  labs(x = "Neighborhood density", y = "Word types (%)") +
  ggtitle(title)+
  theme(axis.text.x = element_text(size=6),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=8))
}
  
plot1 <- plot_pn(mot_on_pp_pn, "Mothers")
plot2 <- plot_pn(bnc_match_ph, "BNC-match")
plot3 <- plot_pn(bnc_all_ph, "BNC-all")

grid.arrange(arrangeGrob(plot1,plot2,plot3,
                         left = textGrob("Word types (%)", rot = 90, vjust = 1),
                         bottom = textGrob("Neighborhood density", rot = 0, vjust = 0.5),
                         nrow = 1))
```

**Figure 5.** Percentage of word types by neighbourhood size, for mothers, BNC-match, and BNC-all. 

<br>

#### Table

Kolmogorov-Smirnov analyses comparing phonotactic probability and neighbourhood density across mothers and BNC-match. 

```{r echo=FALSE, warning=FALSE}
ks.test(mot_on_pp_pn$pp, bnc_match_ph$pp) %>%
  (function(x) {
    tibble(D = x$statistic %>%
             round(2), 
           p = ifelse(x$p.value < 0.001, 
                      "<0.001", 
                      round(x$p.value,2)))
  }) %>%
  cbind(ks.test(mot_on_pp_pn$pn, bnc_match_ph$pn) %>%
          (function(x) {
            tibble(D = x$statistic %>%
                     round(2), 
                   p = ifelse(x$p.value < 0.001, 
                              "<0.001", 
                              round(x$p.value,2)))
  })) %>%
  (function(x) {
    rownames(x) <- "Mothers vs. BNC-match"
    x
  }) %>%
  kable()
  
```

<br>

### Word length

```{r echo=FALSE, warning=FALSE}
ks_phon_len <- ks.test(mot_uni_len$len_phon, bnc_match_len$phon_len) %>%
  (function(x) {
    tibble(D = x$statistic %>%
             round(2),
           p = ifelse(x$p.value < 0.001, "<0.001", round(x$p.value,2)))
  })
```

(see Figure 6, Kolmogorov-Smirnov *D* = `r ks_phon_len$D`, *p* `r ks_phon_len$p`)

```{r len_phon, echo=FALSE, warning=FALSE}
plot_phon_len <- function(df, title) {
  ggplot(df, aes(x=len_phon))+
  geom_bar(aes(y = (..count..)/sum(..count..))) + 
  scale_y_continuous(labels=scales::percent, limits = c(0,0.25)) +
  scale_x_continuous(limits = c(0,16), breaks = seq(0,16,2)) +
  labs(x = "Phonemic length", y = "Word types (%)") +
  ggtitle(title)+
  theme(axis.text.x = element_text(size=6),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=8))
}

plot1 <- plot_phon_len(mot_uni_len, "Mothers")
plot2 <- plot_phon_len(bnc_match_len %>%
                           rename(len_phon = phon_len),
                       "BNC-match")
plot3 <- plot_phon_len(bnc_all_len %>%
                           rename(len_phon = phon_len),
                       "BNC-all")

grid.arrange(arrangeGrob(plot1,plot2,plot3,
                         left = textGrob("Word types (%)", rot = 90, vjust = 1),
                         bottom = textGrob("Phonemic length", rot = 0, vjust = 0.5),
                         nrow = 1))
```

**Figure 6.** Percentage of word types by phonemic length, for mothers, BNC-match, and BNC-all.

<br>

#### Table

95% bootstrapped confidence intervals (*N*=100000) of Kolmogorov-Smirnov statistic (Mothers vs BNC-match) for phonotactic probability, neighborhood density and phonemic length.

```{r echo=FALSE}
KS_CI %>% kable()
```

<br>

### Word frequency

For this measure, to include all types in Spoken BNC (see Table 1. for difference in types between BNC-all-raw and BNC-all) a new random sample from BNC-all-raw was taken and BNC-all-raw was used as full Spoken BNC. 
Word rank was assigned within the **BNC (written + spoken)** and CPWD.

```{r word_freq, echo=FALSE, warning=FALSE, message=FALSE,fig.width=16, fig.height=9}
bncs_mot_comm_raw_fil_1 <- bncs_mot_comm_raw_fil %>%
  select(-freq_cpwd) %>%
  arrange(desc(freq)) %>%
  mutate(rank = 1:nrow(bncs_mot_comm_raw_fil)) %>%
  gather(dataset, rel_freq, c(rel_freq_mot, rel_freq_bnc_sub, rel_freq_bnc)) %>%
  group_by(dataset) %>%
  mutate(rel_freq_cum = cumsum(rel_freq))  %>%
  rename(corpus = dataset)

bncs_mot_comm_raw_fil_2 <- bncs_mot_comm_raw_fil %>%
  select(-freq) %>%
  arrange(desc(freq_cpwd)) %>%
  mutate(rank = 1:nrow(bncs_mot_comm_raw_fil)) %>%
  gather(dataset, rel_freq, c(rel_freq_mot, rel_freq_bnc_sub, rel_freq_bnc)) %>%
  group_by(dataset) %>%
  mutate(rel_freq_cum = cumsum(rel_freq)) %>%
  rename(corpus = dataset)

plot1 <- ggplot(bncs_mot_comm_raw_fil_1, aes(x=log(rank), y=rel_freq_cum, fill=corpus))+
  geom_point(colour="grey")+
  geom_smooth(size=1, colour="black", aes(linetype = corpus)) +
  scale_y_continuous(labels=scales::percent, limits = c(0,1)) +
  ggtitle("BNC")+
  theme(axis.text=element_text(size=18),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=20),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        legend.key.size = unit(1.5, "cm"))+  
  scale_linetype_manual("Corpus", labels = c("BNC-all","BNC-match","Mothers"),values=c("dotted", "longdash", "solid"))+
  guides(fill=FALSE, alpha=F)

plot2 <- ggplot(bncs_mot_comm_raw_fil_2, aes(x=log(rank), y=rel_freq_cum, fill=corpus))+
  geom_point(colour="grey")+
  geom_smooth(size=1, colour="black", aes(linetype = corpus)) +
  scale_y_continuous(labels=scales::percent, limits = c(0,1)) +
  ggtitle("CPWD")+
  theme(axis.text=element_text(size=18),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=20),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        legend.key.size = unit(1.5, "cm"))+  
  scale_linetype_manual("Corpus", labels = c("BNC-all","BNC-match","Mothers"),values=c("dotted", "longdash", "solid"))+
  guides(fill=FALSE, alpha=F)


grid_arrange_shared_legend(plot1, plot2, position = "right", 
                           x_title = "Word rank (log)", y_title = "Cumulative proportion of tokens",
                           x_size = 25, y_size = 25,
                           widths1 = 3.5, widths2 = 0.5, hjust1 = 0.8, vjust1 = 0.3)
```

<br>

A Generalized Additive Model (GAM; nonlinear approach) is used to compare word frequency patterns. In GAM, smooth functions are incorporated of the predictor variables allowing for nonlinear relationships between the covariates and the target variable (see [Clark, 2018](https://m-clark.github.io/generalized-additive-models/)). For this case, the factor variable "corpus" (BNC-match/Mothers) is included in the model allowing to separate smooths for each level; the factor variable is also treated as fixed predictor because the smooths are centered (see [Specifying generalized additive models](https://goo.gl/drrc7H)). The covariate "log(word rank)" is smoothed over. The GAM models considered were specified within the R environment as follows:  

*model = mgvk::gam(tokens_cumulative_proportion ~ corpus + s(log(word_rank), by = corpus), data = BNC/CPWD)*.  

Below we can see the smooth components of the fitted GAM models:

```{r echo = FALSE}
bncs_bnc <-  bncs_mot_comm_raw_fil_1 %>%
  ungroup() %>%
  filter(corpus != "rel_freq_bnc") %>%
  mutate(corpus = factor(corpus)) 

bncs_cpwd <- bncs_mot_comm_raw_fil_2 %>%
  ungroup() %>%
  filter(corpus != "rel_freq_bnc") %>%
  mutate(corpus = factor(corpus))

gam_models <- list(
  mod_bnc1 = gam(rel_freq_cum ~ s(log(rank)), data = bncs_bnc),
  mod_bnc2 = gam(rel_freq_cum ~ corpus + s(log(rank), by = corpus), data = bncs_bnc),
  mod_cpwd1 = gam(rel_freq_cum ~ s(log(rank)), data = bncs_cpwd),
  mod_cpwd2 = gam(rel_freq_cum ~ corpus + s(log(rank), by = corpus), data = bncs_cpwd)
) %>%
  lapply(function(x) list(model = x, summary = summary(x)))
```

```{r echo= FALSE, results='hide', include=FALSE}
plots <- list(
  p1 <- plot(gam_models$mod_bnc2$model),
  p2 <- plot(gam_models$mod_cpwd2$model)
)
```

```{r echo=FALSE}
par(mfrow=c(2,2))
plot(gam_models$mod_bnc2$model, select=1, 
     xlab = "Word rank(log)", ylab = plots[[1]][[1]]$ylab %>%
       str_replace("corpusrel_freq_bnc_sub", "BNC-match"),
     main = "BNC")
plot(gam_models$mod_bnc2$model, select=2, 
     xlab = "Word rank(log)", ylab = plots[[1]][[2]]$ylab %>%
       str_replace("corpusrel_freq_mot", "Mothers"),
     main = "BNC")
plot(gam_models$mod_cpwd2$model, select=1, 
     xlab = "Word rank(log)", ylab = plots[[2]][[1]]$ylab %>%
       str_replace("corpusrel_freq_bnc_sub", "BNC-match"),
     main="CPWD")
plot(gam_models$mod_cpwd2$model, select=2, 
     xlab = "Word rank(log)", ylab = plots[[2]][[2]]$ylab %>%
       str_replace("corpusrel_freq_mot", "Mothers"),
     main="CPWD")
```

<br>

To compare word frequency based on BNC and CPWD, for each model (BNC or CPWD) we compare a base model which contains only the "log(word_rank)" covariate to a model that also includes the factor "corpus" variable. The following tables show different indexes such as *R^2 adjusted*, *GCV* (generalized cross validation score which can be taken as an estimate of the mean square prediction error, used it similar to AIC as a comparative measure to choose among different models, with lower being better), *AIC*, *BIC*, *Deviance* (deviance reduction attributed to this term), *Pr(>Chi)* (Chi-squared test p value). Although we can see from the p values a significant change in the model fit for both BNC and CPWD (related to the large sample size, *N*=`r bncs_bnc %>% nrow()`), we can see that all comparative indexes indicate a greater contribution of the second model in the case of CPWD, capturing the greater difference in word frequency between Mothers and BNC-match when assigning word rank within CPWD. 

<br>

#### Table

```{r echo=FALSE}
gam_mod_table <- tibble(Model = c("BNC [log(rank)]", "BNC [log(rank), corpus]", 
                                  "CPWD [log(rank)]", "CPWD [log(rank), corpus]"),
                        `R^2 (adj)` = c(gam_models %>% sapply(function(x) x$summary$r.sq)),
                        GCV = c(gam_models %>% sapply(function(x) x$summary$sp.criterion))) %>%
  cbind(AIC(gam_models$mod_bnc1$model, gam_models$mod_bnc2$model) %>%
          rbind(AIC(gam_models$mod_cpwd1$model, gam_models$mod_cpwd2$model))) %>%
  cbind(BIC(gam_models$mod_bnc1$model, gam_models$mod_bnc2$model) %>%
          rbind(BIC(gam_models$mod_cpwd1$model, gam_models$mod_cpwd2$model)) %>%
          select(-df)) %>%
  (function(x) {
    row.names(x) <- NULL
    x
  }) 

gam_mod_table %>% kable()
```

<br>

#### Table

```{r echo=FALSE}
gam_mod_table %>%
  select(Model) %>%
  cbind(anova(gam_models$mod_bnc1$model, 
              gam_models$mod_bnc2$model,
              test = "Chisq") %>% as_tibble() %>%
          rbind(anova(gam_models$mod_cpwd1$model, 
                      gam_models$mod_cpwd2$model,
                      test = "Chisq") %>% as_tibble())) %>%
  (function(x) {
    row.names(x) <- NULL
    x
  }) %>%
  mutate(`Pr(>Chi)` = ifelse(`Pr(>Chi)` < 0.001, 0.001, `Pr(>Chi)`)) %>%
  kable()
```
