---
title: "MOT vs Spoken BNC results"
output:
  word_document:
    reference_docx: template.docx
  pdf_document: default

---

Date: `r Sys.Date()`

<br>

```{r echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE, include=FALSE}
lib <- c("tidyverse", "magrittr", "knitr", "gridExtra", "grid")
lapply(lib, require, character.only = TRUE)
load(file="ws_report.RData")
```

```{r echo = FALSE}
grid_arrange_shared_legend <- function(..., nrow = 1,y_title, x_size, y_size,widths1 = 2, widths2=0.2,hjust1=1,vjust1=0, x_title, ncol = length(list(...)), position = c("bottom", "right")) {

  plots <- list(...)
  position <- match.arg(position)
  g <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
  lheight <- sum(legend$height)
  lwidth <- sum(legend$width)
  gl <- lapply(plots, function(x) x + theme(legend.position = "none"))
  gl <- c(gl, nrow = nrow, ncol = ncol)

  combined <- switch(position,
                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
                                            legend,
                                            ncol = 1,
                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
                                           legend,
                                           ncol = 2,
                                           widths = c(widths1, widths2),
                                           left=textGrob(y_title, gp=gpar(fontsize=y_size), rot = 90),
                                           bottom=textGrob(x_title, gp=gpar(fontsize=x_size), vjust = vjust1, hjust = hjust1)))
  grid.newpage()
  grid.draw(combined)

}

round_df <- function(df, digits =2) {
  nums <- vapply(df, is.numeric, FUN.VALUE = logical(1))

  df[,nums] <- round(df[,nums], digits = digits)

  (df)
}
```

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "images/"
)
```

#### Table 1.

```{r echo=FALSE }
BNC_raw_diff <- summary_texts %>%
  (function(x) {
    100 - ((x$Tokens[which(x$Transcription == "BNC-all, Orthographic")] / 
              x$Tokens[which(x$Transcription == "BNC-all-raw, Orthographic")]) * 100)
  }) %>%
  round(2)
```

Total number of orthographic and phonetic tokens and types in mothers and BNC samples, together with Mean Length of Utterance (MLU) information (NOTE. In the table, BNC-all-raw refers to the raw Spoken BNC which includes `r BNC_raw_diff`% of words not available in their phonetic form. Only BNC-all and BNC-match will be used in the following analyses).

```{r echo=FALSE }
summary_texts %>%
  kable()
```

<br>

### Word categories

For word types (see Figure 1 below), plural common nouns and verbs were considered in their root forms ("singular common noun" and "to-infinitive" forms respectively).

<br>

```{r gram_types, echo=FALSE, warning=FALSE}
mot_uni_mor_raw$cat %<>% factor(levels = mot_uni_mor_raw$cat,
                                ordered = T)
bnc_all_ort_types$cat %<>% factor(levels = mot_uni_mor_raw$cat,
                                ordered = T)
bnc_match_ort_types$cat %<>% factor(levels = mot_uni_mor_raw$cat,
                                ordered = T)

plot_gram_types <- function(df, title) {
  ggplot(df, aes(x=cat, y=perc))+
  geom_bar(position="dodge",stat="identity")+
  labs(x = "Grammatical categories", y = "Word types (%)") +
  ggtitle(title)+
  theme(axis.text.x = element_text(size=5,angle = 40, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=8))+
  scale_y_continuous(limits = c(0,62.5))
}

plot1 <- plot_gram_types(mot_uni_mor_raw, "Mothers")
plot2 <- plot_gram_types(bnc_match_ort_types, "BNC-match")
plot3 <- plot_gram_types(bnc_all_ort_types, "BNC-all")

grid.arrange(arrangeGrob(plot1,plot2,plot3,
                         left = textGrob("Word types (%)", rot = 90, vjust = 1),
                         bottom = textGrob("Grammatical categories", rot = 0, vjust = 0.5),
                         nrow = 1))

```

**Figure 1.** Relative word type frequencies of nouns, verbs, adjectives (Adj), adverbs (Adv) and pronouns (Pron) for Mothers, BNC-match and BNC-all. 

<br>

#### Table 2.

Chi-squared analyses comparing the mothers and BNC-match for number of nouns, verbs, adjectives, adverbs and pronouns, for word types (left) and word tokens (right).

```{r echo=FALSE, warning=FALSE}
prop_test <- function(df_mot, df_bnc_match, cat_name, word_type) {
  n_prop_test <- function(df) {
    ( sum(df$freq) / (df$perc %>% sum()) ) * 100
  }
  
  n_mot <- n_prop_test(df_mot)
  n_bnc_match <- n_prop_test(df_bnc_match)
  
  prop_test <- prop.test(x = c(df_mot$freq[which(df_mot$cat == cat_name)], 
                df_bnc_match$freq[which(df_bnc_match$cat == cat_name)]),
                n = c(n_mot, n_bnc_match),
                correct = FALSE)
  
  tibble(!!word_type := cat_name,
         `λ^2` = prop_test$statistic,
         df = prop_test$parameter,
         p = prop_test$p.value)
}

prop_table <- function(df_mot, df_bnc_match, word_type) {
  prop_test(df_mot, df_bnc_match, "NOUN", word_type) %>%
  rbind(prop_test(df_mot, df_bnc_match, "VERB", word_type)) %>%
  rbind(prop_test(df_mot, df_bnc_match, "ADJ", word_type)) %>%
  rbind(prop_test(df_mot, df_bnc_match, "ADV", word_type)) %>%
  rbind(prop_test(df_mot, df_bnc_match, "PRON", word_type)) %>%
  mutate(`λ^2` = round(`λ^2`, 2),
         p = ifelse(p < 0.001, "<0.001", round(p, 2)))
}

prop_table(mot_uni_mor_raw, bnc_match_ort_types, "Word types") %>%
  cbind(prop_table(mot_mor_raw, bnc_match_gram_tokens, "Word tokens")) %>%
  kable()
```


<br>

```{r gram_tokens, echo=FALSE, warning=FALSE}
mot_mor_raw$cat %<>% factor(levels = mot_mor_raw$cat, ordered = T)
bnc_all_gram_tokens$cat %<>% factor(levels = mot_mor_raw$cat, ordered = T)
bnc_match_gram_tokens$cat %<>% factor(levels = mot_mor_raw$cat, ordered = T)

plot_gram_tokens <- function(df, title) {
  ggplot(df, aes(x=cat, y=perc))+
  geom_bar(position="dodge",stat="identity")+
  labs(x = "Grammatical categories", y = "Word tokens (%)") +
  ggtitle(title)+
  theme(axis.text.x = element_text(size=5,angle = 40, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=8))+
  scale_y_continuous(limits = c(0,35))
}
  
plot1 <- plot_gram_tokens(mot_mor_raw, "Mothers")
plot2 <- plot_gram_tokens(bnc_match_gram_tokens, "BNC-match")
plot3 <- plot_gram_tokens(bnc_all_gram_tokens, "BNC-all")

grid.arrange(arrangeGrob(plot1,plot2,plot3,
                         left = textGrob("Word tokens (%)", rot = 90, vjust = 1),
                         bottom = textGrob("Grammatical categories", rot = 0, vjust = 0.5),
                         nrow = 1))

```

**Figure 2.** Relative word token frequencies of verbs, pronouns (Pron), nouns, adverbs (Adv) and adjectives (Adj) for mothers, BNC-match and BNC-all.

<br>

### Utterance-final word categories

```{r gram_last_tokens, echo=FALSE, warning=FALSE}
mot_mor_raw_last$cat %<>% factor(levels = mot_mor_raw_last$cat, ordered = T)
bnc_all_last_gram_tokens$cat %<>% factor(levels = mot_mor_raw_last$cat, ordered = T)
bnc_match_last_gram_tokens$cat %<>% factor(levels = mot_mor_raw_last$cat, ordered = T)

plot_gram_tokens <- function(df, title) {
  ggplot(df, aes(x=cat, y=perc))+
  geom_bar(position="dodge",stat="identity")+
  labs(x = "Grammatical categories", y = "Word tokens (%)") +
  ggtitle(title)+
  theme(axis.text.x = element_text(size=5,angle = 40, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size=8))+
  scale_y_continuous(limits = c(0,35))
}
  
plot1 <- plot_gram_tokens(mot_mor_raw_last, "Mothers")
plot2 <- plot_gram_tokens(bnc_match_last_gram_tokens, "BNC-match")
plot3 <- plot_gram_tokens(bnc_all_last_gram_tokens, "BNC-all")

grid.arrange(arrangeGrob(plot1,plot2,plot3,
                         left = textGrob("Word tokens (%)", rot = 90, vjust = 1),
                         bottom = textGrob("Grammatical categories", rot = 0, vjust = 0.5),
                         nrow = 1))

```

**Figure 3.** Relative word token frequencies of nouns, pronouns (Pron), verbs, adverbs (Adv) and adjectives (Adj) for mothers, BNC-match and BNC-all.

<br>

#### Table 3. 

Chi-squared analyses comparing the mothers and BNC-match for number of nouns, pronouns, verbs, adverbs and adjectives occurring in utterance-final position. 

```{r echo=FALSE, warning=FALSE}
prop_table(mot_mor_raw_last, bnc_match_last_gram_tokens, "Word tokens") %>%
  slice(c(1,5,2,4,3)) %>%
  kable()
```


